---
title: "sfurrr"
output: 
  github_document:
    fig_width: 7.5
    fig_height: 6
  html_preview: false
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "man/figures/", cache = TRUE,
                      warning = FALSE, message = FALSE)
```

<!-- badges: start -->

[![Lifecycle:
experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)

<!-- badges: end -->

An experimental R package to parallelise some functions from the
excellent [{sf}](https://r-spatial.github.io/sf/) and [{rmapshaper}](https://github.com/ateucher/rmapshaper) 
packages using the also brilliant [{furrr}](https://github.com/DavisVaughan/furrr) 
package. Right now, it's just parallel versions of [`st_join` and
`st_filter`](https://r-spatial.github.io/sf/reference/st_join.html) from {sf} and
[`ms_simplify`](https://github.com/ateucher/rmapshaper/blob/master/R/simplify.R) 
from {rmapshaper}. They won't always help and may be slower but, sometimes it might be
useful. This is just messing about right now tbh. The code is deliberately copied 
from {sf}, {rmapshaper} and {furrr} so that it can be used as a drop in replacement.

I've added [{geoarrow}](https://github.com/paleolimbot/geoarrow) as a dependency
to play with using it to pass data between cores - it seems to be fractionally 
faster and solves the issue of passing objects larer than the limit allowed by
furrr... make sure to install {arrow} if you want to try this out.

## Install

    # install.packages("remotes")
    remotes::install_github('h-a-graham/sfurrr')

## Examples...

So here is the some data included in the package - it is the English
cycle network from Open Street Map (downloaded with
[{osmextract}](https://docs.ropensci.org/osmextract/)) and British
counties from [Ordnance
survey](https://osdatahub.os.uk/downloads/open/BoundaryLine). 

```{r show_test_data}
library(sfurrr)

cwe <- cycleways_england()
gbc <-  gb_counties()

basetheme::basetheme("dark")
plot(gbc['geometry'],  axes = TRUE)
plot(cwe['geometry'], add=TRUE, col='#39C17360')


summary(cwe)
summary(gbc)
```

### Spatial Joins

Now, let's say we want to do a spatial join between the cycleways and
the counties so we attach the county data to the cycleway network. This
might allow us to do some summarised stats on the cycle network of
different counties, for example.

So let's do this with {sf} which is loaded by default with {sfurr}.
Let's also get some timings with {tictoc}

```{r intersect_join_sf}
library(tictoc)

tic()
join.sf <-  st_join(cwe,
                  gbc)
toc()

plan(multisession, workers = 4)

tic()
join.sfurr <-  future_st_join(cwe,
                  gbc)
toc()
```

Okay.. so {sf} is actually a lot faster?! Yes, using the simple
st_intersect with a left join (default) is pretty speedy already with
{sf}. But what about more costly spatial operations? Let's try now with
the option `largest=TRUE` which joins based on the largest amount of
intersection.

```{r intersect_largest_join_sf}
# ------------ `st_join` ----------------
tic()
joinL.sf <-  st_join(cwe,
                  gbc, largest=TRUE)
toc()

# ------------ `future_st_join` ----------------
plan(multisession, workers = 8)

tic()
joinL.sfurr <-  future_st_join(cwe,
                  gbc, largest=TRUE)
toc()
```

Okay so now we see that going parallel does indeed offer some potential
uses when using a costly spatial function. 

### Spatial filtering

Once again, here is a comparison of the simplest approach with th
`st_intersect` spatial predicate.

```{r spatial_filter}
# ------------ `st_fiter` ----------------
tic()
filt_t1 <- st_filter(cwe['highway'],
                     gbc[1:50,])
toc()

# ----------- `future_st_filter` -----------------
plan(multisession, workers = 4)
tic()
filt_t2 <- future_st_filter(cwe['highway'],
                            gbc[1:50,])
toc()

```

No surprises, {sf} is faster again. But, what about a more costly
operation. Let's use the `st_within` spatial predicate to filter out
cycleways that are not located entirely within the county areas... This
is kind of pointless for this use case and is just illustrative
really...

```{r spatial_filter_st_within}
# ------------ `st_fiter` ----------------
tic()
within_filt_t1 <- st_filter(joinL.sfurr,
                     gbc[1:50,], .predicate = st_within)
toc()

# ----------- `future_st_filter` -----------------
plan(multisession, workers = 6)

tic()
within_filt_t2 <- future_st_filter(joinL.sfurr,
                            gbc[1:50,], .predicate = st_within)
toc()
```

Cool, so in this case it is faster!

## Summary

This is not a globally useful idea but in some cases, when using very
large spatial datasets, you may get a speed up by running spatial
filters/joins in parallel. Any speed up will depend on the number of
processes you can run; by default, {sfurrr} uses all available cores but
in many cases this is overkill...

```{r final_plot}
plot(gbc['geometry'],  axes = TRUE)
plot(within_filt_t2['Name'], add=TRUE)
```


